---
title: "LAGOS Spatial Analysis"
author: "Mickey Means-Brous"
date: "02/21/22"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
mapviewOptions(fgb = FALSE)
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties

install.packages(c("RApiSerialize", "LAGOSNE", 'USAboundaries'))

##LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus


#load('lake_centers.Rdata')


```



### Convert to spatial data
```{r}
#Look at the column names
names(lake_centers)

#Look at the structure
str(lake_centers)

#View the full dataset
View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()


#Plot all the states to check if they loaded
mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(st_crs(spatial_lakes))

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,] %>%
  mutate(state = 'Minneosta')

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
##Just IL and IA state data
illinoisiowa2 <- states %>%
  filter(name == 'Illinois' | name == 'Iowa') %>%
  st_transform(2163)
mapview(illinoisiowa2)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

There are 16,466 sites in both Illinois and Iowa. This are far more lakes in only Minnesota with 29,038 observations.

```{r}
##Subset LAGOS to IL and IA
illia_lakes <- spatial_lakes[illinoisiowa2,]

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota? Here I want to see a histogram plot with lake size on x-axis and frequency on y axis (check out geom_histogram)

```{r}

iowa <- states %>%
  filter(name == 'Iowa') %>%
  st_transform(st_crs(spatial_lakes))


ia_lakes <- spatial_lakes[iowa,] %>%
  mutate(state = 'Iowa')

both_lakes <- bind_rows(ia_lakes, minnesota_lakes)

ggplot(both_lakes, aes(lake_area_ha)) + 
  geom_histogram() +
  scale_x_log10() +
  facet_wrap(~state)
  
```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares

```{r}
##Lakes by hectares in IL and IA
illia_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 


